A SAT-Based Counterexample Guided Method for Unbounded Synthesis

### 2016-05-08 paper.tex
Changes to section 2 example

\tikzset{every node/.style={solid}}
\tikzstyle{fixed}=[solid]
\begin{figure}
    \centering
    \captionsetup[subfigure]{width=\textwidth,justification=raggedleft}
    \begin{subfigure}[t]{.2\textwidth}
        \centering
        \begin{minipage}[t][3.8cm][t]{\textwidth}
        \begin{tikzpicture}[level distance = 5mm,baseline]
            \node [circle,draw,inner sep=1pt] (root){}
                child {node [circle,draw,inner sep=1pt] {}
                    child {node [circle,draw,inner sep=1pt] {}
                        child {node [circle,draw,inner sep=1pt] {}
                            child {node [circle,draw,inner sep=1pt] {}
                                child {node [circle,draw,inner sep=1pt] {}
                                    node [left=4pt] {\texttt{gr0=0 gr1=0}}
                                }
                                node [left=4pt] {\texttt{req=1}}
                            }
                            node [left=4pt] {\texttt{gr0=0 gr1=0}}
                        }
                        node [left=4pt] {\texttt{req=1}}
                    }
                node [left=4pt] {\texttt{gr0=0 gr1=0}}
                }
                node [left=4pt] {\texttt{req=1}}
                node [above=4pt] {$\langle s, 3 \rangle$};
        \end{tikzpicture}
    \end{minipage}
        \caption{Environment winning trace}
        \label{fig:trace}
    \end{subfigure}
    \begin{subfigure}[t]{.15\textwidth}
        \centering
        \begin{minipage}[t][3.8cm][t]{\textwidth}
        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm,baseline]
            \node [circle,draw] (root){}
                child {node [circle,draw] {}
                    edge from parent [fixed] node [left] {\texttt{req=1}}
                }
                node [left=4pt] {}
                node [above=4pt] {$\langle s, 3 \rangle$};
        \end{tikzpicture}
        \end{minipage}
        \caption{AGT}
        \label{fig:agt}
    \end{subfigure}
    \begin{subfigure}[t]{.2\textwidth}
        \centering
        \begin{minipage}[t][3.8cm][t]{\textwidth}
        \begin{tikzpicture}[level distance = 5mm,baseline]
            \node [circle,draw,inner sep=1pt] (root){}
                child {node [circle,draw,inner sep=1pt] {}
                    child {node [circle,draw,inner sep=1pt] {}
                        child {node [circle,draw,inner sep=1pt] {}
                            child {node [circle,draw,inner sep=1pt] {}
                                child {node [circle,draw,inner sep=1pt] {}
                                    node [left=4pt] {\texttt{gr0=0 gr1=0}}
                                }
                                node [left=4pt] {\texttt{req=1}}
                            }
                            node [left=4pt] {\texttt{gr0=0 gr1=0}}
                        }
                        node [left=4pt] {\texttt{req=1}}
                    }
                node [left=4pt] {\texttt{gr0=1 gr1=1}}
                }
                node [left=4pt] {\texttt{req=1}}
                node [above=4pt] {$\langle s, 3 \rangle$};
        \end{tikzpicture}
        \end{minipage}
        \caption{Controller winning trace}
        \label{fig:trace2}
    \end{subfigure}
    \begin{subfigure}[t]{.15\textwidth}
        \centering
        \begin{minipage}[t][3.8cm][t]{\textwidth}
        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm,baseline]
            \node [circle,draw] (root){}
                child {node [circle,draw] {}
                    child {node [circle,draw] {}
                        % node [left=4pt] {\texttt{gr0=1 gr1=0}}
                        edge from parent [fixed] node [left] {\texttt{req=2}}
                    }
                    % node [left=4pt] {\texttt{gr0=0 gr1=1}}
                    edge from parent [fixed] node [left] {\texttt{req=1}}
                }
%%%                child {node [circle,draw] {}
%%%                    node [left=4pt] {\texttt{req=1}}
%%%                    edge from parent [fixed] node [right] {\texttt{gr=1}}
%%%                }
                node [left=4pt] {}
                node [above=4pt] {$\langle s, 3 \rangle$};
        \end{tikzpicture}
        \end{minipage}
        \caption{AGT}
        \label{fig:agt2}
    \end{subfigure}
    \begin{subfigure}[t]{.2\textwidth}
        \centering
        \begin{minipage}[t][3.8cm][t]{\textwidth}
        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm,baseline]
            \node [circle,draw] (root){}
                child {node [circle,draw] {}
                    child {node [circle,draw] {}
                        child {node [circle,draw] {}
                            edge from parent [fixed] node [right] {\texttt{req=2}}
                        }
                        edge from parent [fixed] node [right] {\texttt{req=2}}
                    }
                    edge from parent [fixed] node [left] {\texttt{req=1}}
                }
                child {node [circle,draw] {}
                    child {node [circle,draw] {}
                        child {node [circle,draw] {}
                            edge from parent [fixed] node [right] {\texttt{req=2}}
                        }
                        edge from parent [fixed] node [right] {\texttt{req=2}}
                    }
                    edge from parent [fixed] node [right] {\texttt{req=2}}
                }
                node [above=4pt] {$\langle s, 3 \rangle$};
        \end{tikzpicture}
        \end{minipage}
        \caption{Refined AGT}
        \label{fig:refined}
    \end{subfigure}
    \caption{Abstract game trees.}
    \label{fig:alltrees}
\end{figure}

@@

    %% Make it clear that this is intuition only
    We present the intuition behind our bounded synthesis method by applying
    its \emph{simplified version} to the running example.
    We begin by finding 
    a trace of length $k$ (here we consider $k=3$) that is winning for the environment, 
    i.e., that starts from the initial state and reaches an
    error state (see Figure~\ref{fig:trace}).  We use a SAT solver to find such a trace,
    precisely as one would do in bounded model checking.
    Given this trace we make an initial
    conjecture that any trace starting with action \texttt{req=1} is winning for the
    environment. This conjecture is captured in the abstract game tree shown in
    Figure~\ref{fig:agt}.  We validate this conjecture by searching for a counterexample
    trace that avoids the error state with the
    first environment action fixed to \texttt{req=1}.   
    Such a trace, that refutes the conjecture, is shown in Figure~\ref{fig:trace2}. 
    In this trace, the controller wins by responding with \texttt{gr0=1, gr1=1} in the first round.
    This move represents controller's partial strategy against the 
    abstract game tree in Figure~\ref{fig:agt}.  This partial strategy is shown in Figure~\ref{}.

    Next we strengthen the abstract game tree taking this partial strategy into account.
    To this end we again use a SAT solver to find a trace where the environment wins while
    the controller plays according to the partial strategy.  In the resulting trace (now shown), 
    the environment plays \texttt{req=2} in the second round.  We refine the abstract game tree
    using this move as shown in Figure~\ref{fig:agt2}.  The controller's partial strategy was to grant two
    resources in the first round, to which the environment responds by now
    making two requests in the second round.

    When the environment cannot refine the tree by extending existing branches, it backtracks 
    and creates new branches.
    Eventually, we obtain the abstract game tree shown in Figure~\ref{fig:refined} for which 
    there does not exist a winning partial strategy on behalf of the controller.  We conclude that 
    the game is winning for the environment.

%    Eventually, the refinements to the environment's candidate will force the
%    controller to grant more resources and result in a game tree such as in
%    Figure~\ref{fig:refined}.

@@

>>>>

\tikzset{every node/.style={solid}}
\tikzstyle{fixed}=[solid]
\begin{figure}
    \centering
    \captionsetup[subfigure]{width=\textwidth,justification=raggedleft}
    \begin{subfigure}[t]{.2\textwidth}
        \centering
        \begin{minipage}[t][3.8cm][t]{\textwidth}
        \begin{tikzpicture}[level distance = 5mm,baseline]
            \node [circle,draw,inner sep=1pt] (root){}
                child {node [circle,draw,inner sep=1pt] {}
                    child {node [circle,draw,inner sep=1pt] {}
                        child {node [circle,draw,inner sep=1pt] {}
                            child {node [circle,draw,inner sep=1pt] {}
                                child {node [circle,draw,inner sep=1pt] {}
                                    node [left=4pt] {\texttt{gr0=0 gr1=0}}
                                }
                                node [left=4pt] {\texttt{req=1}}
                            }
                            node [left=4pt] {\texttt{gr0=0 gr1=0}}
                        }
                        node [left=4pt] {\texttt{req=1}}
                    }
                node [left=4pt] {\texttt{gr0=1 gr1=0}}
                }
                node [left=4pt] {\texttt{req=1}}
                node [above=4pt] {$\langle s, 3 \rangle$};
        \end{tikzpicture}
    \end{minipage}
        \caption{Controller winning trace}
        \label{fig:trace}
    \end{subfigure}
    \begin{subfigure}[t]{.2\textwidth}
        \centering
        \begin{minipage}[t][3.8cm][t]{\textwidth}
        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm,baseline]
            \node [circle,draw] (root){}
                child {node [circle,draw] {}
                    edge from parent [fixed] node [left] {\texttt{gr0=1 gr1=0}}
                }
                node [left=4pt] {}
                node [above=4pt] {$\langle s, 3 \rangle$};
        \end{tikzpicture}
        \end{minipage}
        \caption{AGT}
        \label{fig:agt}
    \end{subfigure}
    \begin{subfigure}[t]{.2\textwidth}
        \centering
        \begin{minipage}[t][3.8cm][t]{\textwidth}
        \begin{tikzpicture}[level distance = 5mm,baseline]
            \node [circle,draw,inner sep=1pt] (root){}
                child {node [circle,draw,inner sep=1pt] {}
                    child {node [circle,draw,inner sep=1pt] {}
                        child {node [circle,draw,inner sep=1pt] {}
                            child {node [circle,draw,inner sep=1pt] {}
                                child {node [circle,draw,inner sep=1pt] {}
                                    node [left=4pt] {\texttt{gr0=0 gr1=0}}
                                }
                                node [left=4pt] {\texttt{req=1}}
                            }
                            node [left=4pt] {\texttt{gr0=0 gr1=0}}
                        }
                        node [left=4pt] {\texttt{req=1}}
                    }
                node [left=4pt] {\texttt{gr0=1 gr1=0}}
                }
                node [left=4pt] {\texttt{req=2}}
                node [above=4pt] {$\langle s, 3 \rangle$};
        \end{tikzpicture}
        \end{minipage}
        \caption{Environment winning trace}
        \label{fig:trace2}
    \end{subfigure}
    \begin{subfigure}[t]{.2\textwidth}
        \centering
        \begin{minipage}[t][3.8cm][t]{\textwidth}
        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm,baseline]
            \node [circle,draw] (root){}
                child {node [circle,draw] {}
                    edge from parent [fixed] node [left] {\texttt{gr0=1 gr1=0}}
                }
                node [left=4pt] {\texttt{req=2}}
                node [above=4pt] {$\langle s, 3 \rangle$};
        \end{tikzpicture}
        \end{minipage}
        \caption{Partial Strategy}
        \label{fig:strategy}
    \end{subfigure}
    \caption{Abstract game trees.}
    \label{fig:alltrees}
\end{figure}

@@

    %% Make it clear that this is intuition only
    We present the intuition behind our bounded synthesis method by applying
    its \emph{simplified version} to the running example.  We begin by finding
    a trace of length $k$ (here we consider $k=3$) that is winning for the
    controller, i.e., that starts from the initial state and avoids the error
    set for three game rounds (see Figure~\ref{fig:trace}).  We use a SAT
    solver to find such a trace, precisely as one would do in bounded model
    checking.  Given this trace we make an initial conjecture that any trace
    starting with action \texttt{gr0=1 gr1=0} is winning for the controller.
    This conjecture is captured in the abstract game tree shown in
    Figure~\ref{fig:agt}.  We validate this conjecture by searching for a
    counterexample trace that reaches an error state with the first controller
    action fixed to \texttt{gr0=1 gr1=0}.   Such a trace, that refutes the
    conjecture, is shown in Figure~\ref{fig:trace2}.  In this trace, the
    environment wins by playing \texttt{req=2} in the first round.  This move
    represents the environment's partial strategy against the abstract game
    tree in Figure~\ref{fig:agt}.  This partial strategy is shown in
    Figure~\ref{fig:strategy}.

    Next we strengthen the abstract game tree taking this partial strategy into account.
    To this end we again use a SAT solver to find a trace where the contoller
    wins while the environment plays according to the partial strategy.  In the
    resulting trace (Figure~\ref{fig:trace3}), the controller plays \texttt{gr0=1 gr1=1} in
    the second round.  We refine the abstract game tree using this move as
    shown in Figure~\ref{fig:refined1}.  The environment's partial strategy was
    to make two requests in the first round, to which the controller responds
    by now granting an additional two resources in the second round.

    When the controller cannot refine the tree by extending existing branches,
    it backtracks and creates new branches. Eventually, we obtain the abstract
    game tree shown in Figure~\ref{fig:refined2} for which there does not exist
    a winning partial strategy on behalf of the environment.  We conclude that
    the bounded game is winning for the controller.

@@

\begin{figure}
    \centering
    \begin{subfigure}[t]{.2\textwidth}
        \centering
        \begin{minipage}[t][3.8cm][t]{\textwidth}
        \begin{tikzpicture}[level distance = 5mm,baseline]
            \node [circle,draw,inner sep=1pt] (root){}
                child {node [circle,draw,inner sep=1pt] {}
                    child {node [circle,draw,inner sep=1pt] {}
                        child {node [circle,draw,inner sep=1pt] {}
                            child {node [circle,draw,inner sep=1pt] {}
                                child {node [circle,draw,inner sep=1pt] {}
                                    node [left=4pt] {\texttt{gr0=0 gr1=0}}
                                }
                                node [left=4pt] {\texttt{req=1}}
                            }
                            node [left=4pt] {\texttt{gr0=1 gr1=1}}
                        }
                        node [left=4pt] {\texttt{req=1}}
                    }
                node [left=4pt] {\texttt{gr0=1 gr1=0}}
                }
                node [left=4pt] {\texttt{req=2}}
                node [above=4pt] {$\langle s, 3 \rangle$};
        \end{tikzpicture}
        \end{minipage}
        \caption{Controller winning trace}
        \label{fig:trace3}
    \end{subfigure}
    \begin{subfigure}[t]{.3\textwidth}
        \centering
        \begin{minipage}[t][3.8cm][t]{\textwidth}
        \hspace*{0.8cm}
        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm,baseline]
            \node [circle,draw] (root){}
                child {node [circle,draw] {}
                    child {node [circle,draw] {}
                        edge from parent [fixed] node [left] {\texttt{gr0=1 gr1=1}}
                    }
                    edge from parent [fixed] node [left] {\texttt{gr0=1 gr1=0}}
                }
                node [above=4pt] {$\langle s, 3 \rangle$};
        \end{tikzpicture}
        \end{minipage}
        \caption{First refined AGT}
        \label{fig:refined1}
    \end{subfigure}
    \begin{subfigure}[t]{.4\textwidth}
        \centering
        \begin{minipage}[t][3.8cm][t]{\textwidth}
        \begin{tikzpicture}[dash pattern = on 2pt off 2pt, level distance = 10mm,baseline]
            \node [circle,draw] (root){}
                child {node [circle,draw] {}
                    child {node [circle,draw] {}
                        child {node [circle,draw] {}
                            edge from parent [fixed] node [left] {\texttt{gr0=1 gr1=1}}
                        }
                        edge from parent [fixed] node [left] {\texttt{gr0=1 gr1=1}}
                    }
                    edge from parent [fixed] node [left] {\texttt{gr0=1 gr1=0}}
                }
                child {node [circle,draw] {}
                    child {node [circle,draw] {}
                        child {node [circle,draw] {}
                            edge from parent [fixed] node [right] {\texttt{gr0=1 gr1=1}}
                        }
                        edge from parent [fixed] node [right] {\texttt{gr0=1 gr1=1}}
                    }
                    edge from parent [fixed] node [right] {\texttt{gr0=1 gr1=1}}
                }
                node [above=4pt] {$\langle s, 3 \rangle$};
        \end{tikzpicture}
        \end{minipage}
        \caption{Final Refined AGT}
        \label{fig:refined2}
    \end{subfigure}
    \caption{Refined abstract game trees.}
    \label{fig:refinedtrees}
\end{figure}
